---
title: "S2nd_compositescorePlot_scatterplot"
author: "Xiaoyu xu"
date: "4/17/2024"
output: html_document
---

This script generate the schema of fluid cognition structure and scatter plots between SC strength and fluid cognition for 3 exemplified connections.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(mgcv)
library(parallel)
library(corrplot)
library(ggcorrplot)
library(ggplot2)
library(RColorBrewer)
library(tidyverse)
library(geomtextpath)
interfileFolder <- '/Users/xuxiaoyu_work/Cuilab/DMRI_network_development/SC_development/interdataFolder_ABCD'
FigureFolder<-'/Users/xuxiaoyu_work/Cuilab/DMRI_network_development/SC_development/Figure_ABCD_final/SA12/CV75/cognition'
resultFolder<-'/Users/xuxiaoyu_work/Cuilab/DMRI_network_development/SC_development/results_ABCD'
functionFolder<-'/Users/xuxiaoyu_work/Cuilab/DMRI_network_development/SC_development/Rcode_SCdevelopment/gamfunction'
source(paste0(functionFolder, "/gamcog.R"))
## load data
SCdata.all<-readRDS(paste0(interfileFolder, '/SCdata_SA12_CV75_sumSCinvnode.sum.msmtcsd.combatNTBfluid.rds'))
SCdata <- SCdata.all
SA12_10 <- read.csv(paste0(interfileFolder, '/SA12_10.csv'))
SA12_10 <- SA12_10[order(SA12_10$SCrank),]

```

## 1.fluid score component, Figure 6a

```{r ComponentPlot}
# flanker, DCCS, PSM, lswm, pattern
fluidplot <- data.frame(
  category = c("Inhibition", "Flexibility", "Episodic memory", "Working memory", "Pattern comparison"),
  value = c(10, 10, 10, 10, 10)
)
fluidplot$category <- as.factor(fluidplot$category)
fluidplot$x <- rep(0.5,5)
fluidplot$angle = c(0, 108,  216, -216, -108)
fluidplot$angle[2:5] = fluidplot$angle[2:5]+180
ggplot(fluidplot, aes(x = x, y = value, fill = category)) +
  geom_bar(stat = "identity", width = 1, show.legend=F) +
  coord_polar(theta = "y") +scale_fill_manual(values=brewer.pal(5, "Set3"))+
  xlim(-1, 1)+geom_text(aes(label = category, angle=angle), position = position_stack(vjust = 0.5),size=7)+
  annotate("text", x=-1, y=0, label="Fluid Cognition",size=7)+
  theme_void()

ggsave(paste0(FigureFolder, '/NTB_fluidscore_component.svg'), width = 16, height=16, units = "cm")

```

## 2. cognition ~ SC strength scatter plot, Figure 6d
Residuals regressed out age, gender, race, handedness and head motions were used in these plots. Three connections located at different positions along the connectional axis were selected for exemples.

```{r scatterplots}
Cogvar <- "nihtbx_fluidcomp_uncorrected"
CVthr <- 75
SC_Cog_results.df <- readRDS(paste0(interfileFolder, "/SC_Cog_results_", Cogvar,"_CV", CVthr,".rds"))
# plot data
nonna_index<-which(!is.na(SCdata[ ,Cogvar]))
SCdata.cog<-SCdata[nonna_index,]
SCdata.cog <- SCdata.cog[SCdata.cog$eventname=="baseline_year_1_arm_1",]
SCdata.cog[,str_detect(names(SCdata.cog), "SC.")] <- lapply(SCdata.cog[,str_detect(names(SCdata.cog), "SC.")], as.numeric)
dataname<-"SCdata.cog"
smooth_var<-"age"
covariates<-"gender+handness+race+mean_fd"
knots<-3
corrmethod<-"pearson"
plotdata <- mclapply(1:78, function(x) {
  SClabel <- grep("SC.", names(SCdata.cog), value=T)[x]
  region <- SClabel
  gamresult.df <-
    gam.fit.cognition(
      Cogvar,
      dataname,
      region, 
      smooth_var, 
      covariates, 
      knots,
      corrmethod, set_fx = TRUE, stats_only = FALSE)
  gamresult.df <- as.data.frame(gamresult.df[[2]])
  return(gamresult.df)
}, mc.cores = 4)

SC_Cog_results.df$SClabelorder <- c(1:78)
SC_Cog_results.df <- merge(SC_Cog_results.df, SA12_10, by.x="cognition_var", by.y="SC_label")
SC_Cog_results.df <- SC_Cog_results.df[order(SC_Cog_results.df$SClabelorder),]
#N=22, 44, 74
N=22
plotdata_N <- plotdata[[N]]
SCrank = SC_Cog_results.df$SCrank[N]
print(paste("The figure below is for connection with a rank of", SCrank))
ggplot(data = plotdata_N)+
  geom_point(aes(x=SCres, y=cogres), color="grey", size=0.8)+
  geom_smooth(aes(x=SCres, y=cogres),linewidth=1.4, method="lm", color="black")+
  labs(x="SC strength", y="Fluid cognition")+
  theme_classic()+
  theme(axis.text=element_text(size=22, color="black"), 
        axis.title =element_text(size=22),aspect.ratio = 0.8,
        axis.line = element_line(linewidth = 0.5),axis.ticks = element_line(linewidth = 0.5),
        plot.title = element_text(size=20, hjust = 0.5, vjust=0),
        plot.subtitle = element_text(size=15, hjust = 0.9, vjust=-6),
        plot.background=element_rect(fill="transparent"),
        panel.background=element_rect(fill="transparent"),
        legend.position = "none")

ggsave(paste0(FigureFolder, "/", Cogvar, "_rev/SC", N, "_scatterplot.svg"), width=14, height=12, units = "cm")
ggsave(paste0(FigureFolder, "/", Cogvar, "_rev/SC", N, "_scatterplot.tiff"), width=14, height=14, units = "cm")

N=44
plotdata_N <- plotdata[[N]]
SCrank = SC_Cog_results.df$SCrank[N]
print(paste("The figure below is for connection with a rank of", SCrank))
ggplot(data = plotdata_N)+
  geom_point(aes(x=SCres, y=cogres), color="grey", size=0.8)+
  geom_smooth(aes(x=SCres, y=cogres),linewidth=1.4, method="lm", color="black")+
  labs(x="SC strength", y="Fluid cognition")+
  theme_classic()+
  theme(axis.text=element_text(size=22, color="black"), 
        axis.title =element_text(size=22),aspect.ratio = 0.8,
        axis.line = element_line(linewidth = 0.5),axis.ticks = element_line(linewidth = 0.5),
        plot.title = element_text(size=20, hjust = 0.5, vjust=0),
        plot.subtitle = element_text(size=15, hjust = 0.9, vjust=-6),
        plot.background=element_rect(fill="transparent"),
        panel.background=element_rect(fill="transparent"),
        legend.position = "none")

ggsave(paste0(FigureFolder, "/", Cogvar, "_rev/SC", N, "_scatterplot.svg"), width=14, height=12, units = "cm")
ggsave(paste0(FigureFolder, "/", Cogvar, "_rev/SC", N, "_scatterplot.tiff"), width=14, height=14, units = "cm")

N=74
plotdata_N <- plotdata[[N]]
SCrank = SC_Cog_results.df$SCrank[N]
print(paste("The figure below is for connection with a rank of", SCrank))
ggplot(data = plotdata_N)+
  geom_point(aes(x=SCres, y=cogres), color="grey", size=0.8)+
  geom_smooth(aes(x=SCres, y=cogres),linewidth=1.4, method="lm", color="black")+
  labs(x="SC strength", y="Fluid cognition")+
  theme_classic()+
  theme(axis.text=element_text(size=22, color="black"), 
        axis.title =element_text(size=22),aspect.ratio = 0.8,
        axis.line = element_line(linewidth = 0.5),axis.ticks = element_line(linewidth = 0.5),
        plot.title = element_text(size=20, hjust = 0.5, vjust=0),
        plot.subtitle = element_text(size=15, hjust = 0.9, vjust=-6),
        plot.background=element_rect(fill="transparent"),
        panel.background=element_rect(fill="transparent"),
        legend.position = "none")

ggsave(paste0(FigureFolder, "/", Cogvar, "_rev/SC", N, "_scatterplot.svg"), width=14, height=12, units = "cm")
ggsave(paste0(FigureFolder, "/", Cogvar, "_rev/SC", N, "_scatterplot.tiff"), width=14, height=14, units = "cm")
```

