---
title: "abcd_SIEMENS_dMRI_datamerge"
author: "Xiaoyu xu"
date: "3/6/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(psych)
library(openxlsx)

info_path <- "/Users/xuxiaoyu_work/Cuilab/open_dataset_information/ABCD/info"
datapath <- "/Volumes/exchange/abcd-data-release-5.0/abcd-data-release-5.1/core"
datapath4 <- "/Users/xuxiaoyu_work/Cuilab/DMRI_network_development/ABCD_download_organise/behavior"
```

## Screen data

This script is to screen the data from data-screening  based on demographic and diagnosis.

## 1. screen data by MRI information
Because all the MRI data were downloaded and organised according to files from release 4.0, the imaging data will be screened according to files from release 4.0.
```{r MRI_screen}
abcd_mri01 <- read.xlsx(paste0(datapath4, '/abcd_mri01.xlsx'))
abcd_mri01 <- abcd_mri01[-1,]
abcd_mri01 <- abcd_mri01[,c(5,9,12, 14)]
abcd_imgincl01 <- read.xlsx(paste0(datapath4, '/abcd_imgincl01.xlsx'))
abcd_imgincl01 <- abcd_imgincl01[-1,-c(1:4, 6:8, 10,18)]
data_raw <- read.csv(paste0(info_path, '/data_raw.csv')) # successfully downloaded raw data
data_nifti <- read.csv(paste0(info_path, '/SIEMENS_niftilist.csv')) # successfully converted to nifti
failBIDScheck <- read.table(paste0(info_path, '/dmrismridcmnum.txt'), header = T)
names(failBIDScheck)[4:5] <- c("dmridcmnum", "smridcmnum")
failBIDScheck <- failBIDScheck %>% filter(dmridcmnum!=103 | smridcmnum!=176)
# If failed unzip, DCM number will be 0; if image is incomplete, DCM number won't be equal to 103 or 176.
failBIDScheck$scanID <- paste0(failBIDScheck$subID, "_ses-", failBIDScheck$eventname)
failBIDScheck <- failBIDScheck %>% distinct(scanID, .keep_all = T)
process_df <- read.csv(paste0(info_path, '/qsi_S_addinfo.csv'))
process_df <- process_df[,c("scanID", "subID", "qsiprepfinished", "failT1Parcellation")]

demodf_s1 <- abcd_mri01
demodf_s1 <- merge(demodf_s1, abcd_imgincl01, by=c("src_subject_id", "eventname"))
#1 only images from SIEMENS scanner
demodf_s1 <- demodf_s1 %>% filter(mri_info_manufacturer=="SIEMENS")
num_SIEMENS <- nrow(demodf_s1)
print(paste0("1st, we only include scans from SIEMENS scanner, ", nrow(demodf_s1), " were included, ", table(demodf_s1$eventname)[2], " from baseline; ", table(demodf_s1$eventname)[1], " from 2year-FU; ", length(unique(demodf_s1$src_subject_id)), " unique participants."))
#2 data successfully downloaded & have all 3 types of images
data_raw$eventname <- gsub("b", "baseline_year_1_arm_1", data_raw$eventname)
data_raw$eventname <- gsub("2", "2_year_follow_up_y_arm_1", data_raw$eventname)
data_raw$subID <- paste0("sub-NDAR", data_raw$subID)
demodf_s1$subID <- paste0("sub-", gsub("_", "", demodf_s1$src_subject_id))
demodf_s1 <- merge(demodf_s1, data_raw, by=c("subID", "eventname"), all.x = T)
ndownload_S <- length(which(demodf_s1$num_dMRI_raw >0 | demodf_s1$num_sMRI_raw >0 | demodf_s1$num_fmap_AP_raw >0))
demodf_s1 <- demodf_s1 %>% filter(num_dMRI_raw>0 & num_sMRI_raw>0 & num_fmap_AP_raw>0)
num_download <- nrow(demodf_s1)
print(paste0("2nd, we successfully downloaded ", num_download, " scans with dMRI, sMRI and fmap.", table(demodf_s1$eventname)[2], " from baseline; ", table(demodf_s1$eventname)[1], " from 2year-FU; ", length(unique(demodf_s1$src_subject_id)), " unique participants."))

#3 not recommended according to official QC (criteria referred to release 4.0 notes)
num_base_2year_now <- table(demodf_s1$eventname)
demodf_s1 <- demodf_s1 %>% filter(imgincl_t1w_include==1 & imgincl_dmri_include==1)
num_base_2year_after <- table(demodf_s1$eventname)
num_base_2year_diff <- num_base_2year_now - num_base_2year_after
num_exclu4qc <- num_download-nrow(demodf_s1)
print(paste0("3rd, we exclude ", num_exclu4qc, " scans for no recommendation from official QC.", num_base_2year_diff[2], " from baseline, ", num_base_2year_diff[1], " from 2-year FU."))
#4 successful BIDS
data_nifti$scanID <- paste0(data_nifti$subID, '_', data_nifti$eventname)
SIEMENS_nifti_complete <- data.frame(scanID=unique(data_nifti$scanID), dMRI=rep(NA, times=length(unique(data_nifti$scanID))),
                                     sMRI=rep(NA, times=length(unique(data_nifti$scanID))), fmap_AP=rep(NA, times=length(unique(data_nifti$scanID))),
                                     fmap_PA=rep(NA, times=length(unique(data_nifti$scanID))))
for (i in 1:nrow(SIEMENS_nifti_complete)){
  scanID <- SIEMENS_nifti_complete$scanID[i]
  indexID <- which(data_nifti$scanID==scanID)
  data.tmp <- data_nifti$filename[indexID]
  dwinum <- length(which(str_detect(data.tmp, "_dwi")))
  sMRInum <- length(which(str_detect(data.tmp, "_T1w")))
  fmap_APnum <- length(which(str_detect(data.tmp, "_dir-AP") | str_detect(data.tmp, "FM-AP")))
  fmap_PAnum <- length(which(str_detect(data.tmp, "_dir-PA") | str_detect(data.tmp, "FM-PA")))
  SIEMENS_nifti_complete[i, 2:5] <- c(dwinum, sMRInum, fmap_APnum, fmap_PAnum)
  SIEMENS_nifti_complete$eventnames[i] <- data_nifti$eventname[indexID[1]]
}
SIEMENS_nifti_complete$completeness <- (SIEMENS_nifti_complete$dMRI>0 & SIEMENS_nifti_complete$sMRI>0 & SIEMENS_nifti_complete$fmap_AP>0)
demodf_s1$eventname2 <- str_replace_all(demodf_s1$eventname, "_(\\w)", function(x) toupper(x))
demodf_s1$scanID <- paste0(demodf_s1$subID, "_ses-", gsub("_", "", demodf_s1$eventname2))
demodf_s1 <- merge(demodf_s1, SIEMENS_nifti_complete, by="scanID", all.x=T)
demodf_s1$completeness[which(is.na(demodf_s1$completeness))] <-0

scanlist_noBIDS <- demodf_s1$scanID[which(demodf_s1$completeness!=1)]
failBIDScheck1 <- failBIDScheck[failBIDScheck$scanID %in% scanlist_noBIDS,]

num_incompletedata <- length(which((failBIDScheck1$dmridcmnum>0 & failBIDScheck1$dmridcmnum<103) | (failBIDScheck1$smridcmnum>0 & failBIDScheck1$smridcmnum<176)))
num_failzip <- length(scanlist_noBIDS) - num_incompletedata

num_base_2year_now <- table(demodf_s1$eventname)
demodf_s1 <- demodf_s1 %>% filter(completeness==1)

demodf_s1 <- merge(demodf_s1, process_df, by="scanID", all.x = T)
scanlist_noBIDS2 <- demodf_s1$scanID[is.na(demodf_s1$qsiprepfinished)]
failBIDScheck2 <- failBIDScheck[failBIDScheck$scanID %in% scanlist_noBIDS2,]
scanlist_noBIDS2[!scanlist_noBIDS2 %in% failBIDScheck2$scanID] # failed dcm2nii
num_incompletedata <- num_incompletedata+length(which((failBIDScheck2$dmridcmnum>0 & failBIDScheck2$dmridcmnum<103) | (failBIDScheck2$smridcmnum>0 & failBIDScheck2$smridcmnum<176)))
num_failzip <- num_failzip+nrow(failBIDScheck2) - length(which((failBIDScheck2$dmridcmnum>0 & failBIDScheck2$dmridcmnum<103) | (failBIDScheck2$smridcmnum>0 & failBIDScheck2$smridcmnum<176)))
num_faildcm2nii <- length(scanlist_noBIDS2) - nrow(failBIDScheck2)
demodf_s1 <- demodf_s1[!is.na(demodf_s1$qsiprepfinished),]
num_base_2year_after <- table(demodf_s1$eventname)
num_base_2year_diff <- num_base_2year_now - num_base_2year_after
for (i in 1){
  print(paste0("4th, ", nrow(demodf_s1), "scans were successfully organised to BIDS, ", table(demodf_s1$eventname)[2], " from baseline, ",  table(demodf_s1$eventname)[1], " from 2year-Fu, ", length(unique(demodf_s1$src_subject_id)), " unique participants."))
print(paste0("Appendex to 4th: ", num_incompletedata, " scans failed for incomplete data; ", num_failzip, " scans failed unzip; ", num_faildcm2nii, " scans failed dcm2niix.", num_base_2year_diff[2], " from baseline, ", num_base_2year_diff[1], " from 2-year FU."))
}

```

## 2. screen data by demographic information
```{r}
demodf <- read.csv(paste0(info_path, '/demo_sublist_criteria.csv'))
demodf <- demodf %>%
  mutate(race_ethnicity = if_else(eventname == "baseline_year_1_arm_1" & is.na(race_ethnicity), 0, race_ethnicity))
demodf <- demodf %>% group_by(src_subject_id) %>% 
  mutate(race_ethnicity = zoo::na.locf(race_ethnicity)) %>%
  mutate(race_ethnicity = zoo::na.locf(race_ethnicity, fromLast = TRUE, na.rm = F))
demodf$race_ethnicity <- as.factor(demodf$race_ethnicity)
summary(demodf$race_ethnicity)
# 1 = White; 2 = Black; 3 = Hispanic; 4 = Asian; 5 = Other; 0=NA
abcd_y_lt <- read.csv(paste0(datapath, '/abcd-general/abcd_y_lt.csv'))
abcd_y_lt <- abcd_y_lt[which(abcd_y_lt$eventname=='baseline_year_1_arm_1' | abcd_y_lt$eventname=='2_year_follow_up_y_arm_1'),]
abcd_y_lt <- abcd_y_lt[,c(1:3,9)]
nc_y_ehis <- read.csv(paste0(datapath, '/neurocognition/nc_y_ehis.csv'))
nc_y_ehis <- nc_y_ehis[nc_y_ehis$eventname=='baseline_year_1_arm_1',]
nc_y_ehis <- nc_y_ehis[,c(1,8)]

demodf <- merge(demodf, abcd_y_lt, by=c("src_subject_id", "eventname"))
demodf <- merge(demodf, nc_y_ehis, by=c("src_subject_id"), all.x = T)
demodf_s2 <- merge(demodf_s1, demodf, by=c("src_subject_id", "eventname"), all.x = T)
#1
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(if_language_y==FALSE & if_language_p==FALSE)
num_exclu4language = nrow(demodf_s1)-nrow(demodf_s2)
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4language <- num_base_2year_now - num_base_2year_after
#2
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(if_TBI==FALSE)
num_exclu4TBI = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4TBI <- num_base_2year_now - num_base_2year_after
#3
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(if_sensory==FALSE)
num_exclu4sensory = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language-num_exclu4TBI
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4sensory <- num_base_2year_now - num_base_2year_after
#4
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(if_intel==FALSE)
num_exclu4inel = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language-num_exclu4TBI-num_exclu4sensory
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4inel <- num_base_2year_now - num_base_2year_after
#5
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(if_medical==FALSE)
num_exclu4medical = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language-num_exclu4TBI-num_exclu4sensory-num_exclu4inel
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4medical <- num_base_2year_now - num_base_2year_after
#6
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(if_contra==FALSE)
num_exclu4contra = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language-num_exclu4TBI-num_exclu4sensory-num_exclu4inel-num_exclu4medical
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4contra <- num_base_2year_now - num_base_2year_after
#7
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(if_premature==FALSE)
num_exclu4premature = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language-num_exclu4TBI-num_exclu4sensory-num_exclu4inel-num_exclu4medical-num_exclu4contra
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4premature <- num_base_2year_now - num_base_2year_after
#8
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% filter(!is.na(interview_age) & if_gender==F)
num_exclu4agegender = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language-num_exclu4TBI-num_exclu4sensory-num_exclu4inel-num_exclu4medical-num_exclu4contra-num_exclu4premature
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4agegender <- num_base_2year_now - num_base_2year_after
#9
num_base_2year_now <- table(demodf_s2$eventname)
demodf_s2 <- demodf_s2 %>% drop_na(c("ehi_y_ss_scoreb"))
num_exclu4hand = nrow(demodf_s1)-nrow(demodf_s2)-num_exclu4language-num_exclu4TBI-num_exclu4sensory-num_exclu4inel-num_exclu4medical-num_exclu4contra-num_exclu4premature-num_exclu4agegender
num_base_2year_after <- table(demodf_s2$eventname)
num_base_2year_diff4hand <- num_base_2year_now - num_base_2year_after

demodf_s2[,6:23] <- NULL
write.csv(demodf_s2, paste0(info_path, '/DemodfScreenByDemoinfo.csv'), row.names = F)
for (i in 1){
  print(paste0("1st, we exclude ", num_exclu4language, " scans for English proficiency in children or English/Spanish proficiency in their parents.", num_base_2year_diff4language[2], " from baseline, ", num_base_2year_diff4language[1], " from 2-year FU."))
  print(paste0("2nd, we exclude ", num_exclu4TBI, " scans for mild and above TBI.", num_base_2year_diff4TBI[2], " from baseline, ", num_base_2year_diff4TBI[1], " from 2-year FU."))
  print(paste0("3rd, we exclude ", num_exclu4sensory, " scans for sensory issues.", num_base_2year_diff4sensory[2], " from baseline, ", num_base_2year_diff4sensory[1], " from 2-year FU."))
  print(paste0("4th, we exclude ", num_exclu4inel, " scans for intellectual disability.", num_base_2year_diff4inel[2], " from baseline, ", num_base_2year_diff4inel[1], " from 2-year FU."))
  print(paste0("5th, we exclude ", num_exclu4medical, " scans for major medical or neurological conditions.", num_base_2year_diff4medical[2], " from baseline, ", num_base_2year_diff4medical[1], " from 2-year FU."))
  print(paste0("6th, we exclude ", num_exclu4contra, " scans for contraindications to MRI scanning.", num_base_2year_diff4contra[2], " from baseline, ", num_base_2year_diff4contra[1], " from 2-year FU."))
  print(paste0("7th, we exclude ", num_exclu4premature, " scans for premature or low birth weight.", num_base_2year_diff4premature[2], " from baseline, ", num_base_2year_diff4premature[1], " from 2-year FU."))
  print(paste0("8th, we exclude ", num_exclu4agegender, " scans for invalid age & sex.", num_base_2year_diff4agegender[2], " from baseline, ", num_base_2year_diff4agegender[1], " from 2-year FU."))
  print(paste0("9th, we exclude ", num_exclu4hand, " scans for invalid handedness.", num_base_2year_diff4hand[2], " from baseline, ", num_base_2year_diff4hand[1], " from 2-year FU."))
  print(paste0("Finally, ", nrow(demodf_s2), " scans left after this step, ", table(demodf_s2$eventname)[2], " from baseline, ",  table(demodf_s2$eventname)[1], " from 2year-Fu, ", length(unique(demodf_s2$src_subject_id)), " unique participants."))
}

```

## 3. data successfully passed qsiprep & head motion criteria

```{r}
demodf_s3 <- demodf_s2
# 1 qsiprep finished
num_base_2year_now <- table(demodf_s3$eventname)
num_failqsiprep <- length(which(demodf_s3$qsiprepfinished==FALSE | demodf_s3$failT1Parcellation==TRUE))
demodf_s3 <- demodf_s3[which(demodf_s3$qsiprepfinished==TRUE & demodf_s3$failT1Parcellation==FALSE), ]
num_base_2year_after <- table(demodf_s3$eventname)
num_base_2year_diff4failqsiprep <- num_base_2year_now - num_base_2year_after

demodf_orig <- read.csv(paste0(info_path, '/qsi_S_addinfo.csv'))
# 2 high head motion
demodf_s3$scanID.x <- demodf_s3$subID.x <- demodf_s3$subID.y <- NULL
demodf_s3 <- dplyr::rename(demodf_s3, scanID="scanID.y")
demodf_s3 <- demodf_s3 %>% left_join(select(demodf_orig, scanID, mean_fd), by="scanID")
highmotion_thr <- mean(demodf_s3$mean_fd)+3*sd(demodf_s3$mean_fd)
demodf_s3 <- demodf_s3 %>% filter(mean_fd < highmotion_thr)
num_exclu4highmotion <- nrow(demodf_s2) - num_failqsiprep - nrow(demodf_s3)
# 2. sites with fewer scans
tab <- table(demodf_s3$site_id_l)
siteexclude <- names(tab)[which(tab < 100)]
# clean
demodf_s3$subID.y <- NULL
demodf_s3[,which(str_detect(names(demodf_s3),"_ft") | str_detect(names(demodf_s3), "_raw") | str_detect(names(demodf_s3), "if_"))] <- NULL
#names(demodf_s3)
demodf_s3$eventname2 <- demodf_s3$eventnames <- NULL
abcd_p_demo <- read.csv(paste0(datapath,'/abcd-general/abcd_p_demo.csv'))

abcd_p_demo <- abcd_p_demo %>% group_by(src_subject_id) %>% 
  mutate(demo_sex_v2 = zoo::na.locf(demo_sex_v2)) %>%
  mutate(demo_sex_v2 = zoo::na.locf(demo_sex_v2, fromLast = TRUE, na.rm = FALSE)) 
demodf_s3 <- demodf_s3 %>% left_join(select(abcd_p_demo, c("src_subject_id", "eventname", "demo_sex_v2")), join_by("src_subject_id", "eventname"))

demodf_s3 <- demodf_s3 %>% dplyr::rename(c(siteID="site_id_l",age="interview_age",sex="demo_sex_v2", handness="ehi_y_ss_scoreb"))
write.csv(demodf_s3, paste0(info_path, '/DemodfScreenFinal.csv'), row.names = F)
for (i in 1){
  print(paste0("1st, ", num_failqsiprep, " failed processing. Structural connectomes were successfully constructed for ", sum(num_base_2year_after), " scans.", num_base_2year_after[2], " from baseline; ", num_base_2year_after[1], " from 2year-FU; "))
  print(paste0("2nd, we exclude ", num_exclu4highmotion, " scans for high head motion."))
  print(paste0("3rd, scans from ", length(siteexclude), " site were fewer than 100."))
  print(paste0("Finally, we included ", nrow(demodf_s3), " scans, ", table(demodf_s3$eventname)[2], " from baseline, ", table(demodf_s3$eventname)[1], " from 2year-FU, ", length(unique(demodf_s3$src_subject_id)), " unique participants."))
}

```

## 4. merge siteID from release4 & cognition & diagnosis
```{r}
# diagnosis
diagnosis_subinfo <- read.csv(paste0(info_path, '/diagnosis_subinfo_l.csv'))
diagnosis_subinfo_consist <- read.csv(paste0(info_path, '/diagnosis_subinfo_consist_l.csv'))
demodf_s3 <- demodf_s3 %>% left_join(select(diagnosis_subinfo,scanID, adhd, anx, odd, conduct, dep, td), by="scanID")
demodf_s3 <- demodf_s3 %>% left_join(diagnosis_subinfo_consist, by="subID")
disorder_type <- c("adhd", "anx", "odd", "dep", "td")
write.csv(demodf_s3, paste0(info_path, '/DemodfScreenFinal.csv'), row.names = F)
for (disorder_now in disorder_type){
  num_scans <- table(demodf_s3[,disorder_now])[2]
  num_consist <- table(demodf_s3[,disorder_now],demodf_s3[,paste0(disorder_now, "_consistent")])[2,2]
  print(paste0("In the final sample, ",num_scans, " scans are diagnosed as ", disorder_now, ", ", num_consist, " scans are consistent at two visits."))
}

# cognition
nc_y_nihtb <- read.csv(paste0(datapath, '/neurocognition/nc_y_nihtb.csv'))
interest_cogvar <- c("src_subject_id", "eventname","nihtbx_flanker_uncorrected", "nihtbx_list_uncorrected", "nihtbx_cardsort_uncorrected", "nihtbx_pattern_uncorrected", "nihtbx_fluidcomp_uncorrected", "nihtbx_cryst_uncorrected", "nihtbx_totalcomp_uncorrected")
demodf_s3 <- demodf_s3 %>% left_join(select(nc_y_nihtb, all_of(interest_cogvar)), by=join_by(src_subject_id, eventname))
print(paste0(table(demodf_s3$eventname[!is.na(demodf_s3$nihtbx_fluidcomp_uncorrected)])[2], " participants in cognition analyses."))
# adhd & odd
data_comorbidity <- demodf_s3[,disorder_type[c(1,3)]]
idx_comorbidity <- which(rowSums(data_comorbidity) > 1)
demodf_s3$comorbidity_adhd_odd <- (rowSums(data_comorbidity) > 1)
demodf_s3_cog <- demodf_s3 %>% drop_na(nihtbx_fluidcomp_uncorrected)
demodf_s3_cog <- demodf_s3_cog[str_detect(demodf_s3_cog$eventname, "base"),]
for (disorder_now in disorder_type[c(1,3)]) 
{
  idx_disorder <- which(demodf_s3[,disorder_now]==1 & demodf_s3$comorbidity_adhd_odd==FALSE)
  #print(paste0(disorder_now,' (exclude comorbidity): ', as.character(length(idx_disorder)), ' scans' ))
  comorbidity_dis <- demodf_s3[,disorder_now] * !demodf_s3$comorbidity_adhd_odd
  num_consist <- table(comorbidity_dis,demodf_s3[,paste0(disorder_now, "_consistent")])[2,2]
  print(paste0("In the final sample, ",length(idx_disorder), " scans are diagnosed as ", disorder_now, " (exclude comorbidity odd & adhd), ", num_consist, " scans are consistent at two visits."))
  
  idx_disorder <- which(demodf_s3_cog[,disorder_now]==1 & demodf_s3_cog$comorbidity_adhd_odd==FALSE)
  #print(paste0(disorder_now,' (exclude comorbidity): ', as.character(length(idx_disorder)), ' scans' ))
  comorbidity_dis <- demodf_s3_cog[,disorder_now] * !demodf_s3_cog$comorbidity_adhd_odd
  num_consist <- table(comorbidity_dis,demodf_s3_cog[,paste0(disorder_now, "_consistent")])[2,2]
  print(paste0("In the cognition sample, ",length(idx_disorder), " scans are diagnosed as ", disorder_now, " (exclude comorbidity  odd & adhd), ", num_consist, " scans are consistent at two visits."))
}

# adhd & anx & odd
data_comorbidity <- demodf_s3[,disorder_type[c(1,2,3)]]
demodf_s3$comorbidity_adhd_odd_anx <- (rowSums(data_comorbidity) > 1)
demodf_s3_cog <- demodf_s3 %>% drop_na(nihtbx_fluidcomp_uncorrected)
demodf_s3_cog <- demodf_s3_cog[str_detect(demodf_s3_cog$eventname, "base"),]
for (disorder_now in disorder_type[c(1,2,3)]) 
{
  idx_disorder <- which(demodf_s3[,disorder_now]==1 & demodf_s3$comorbidity_adhd_odd_anx==FALSE)
  #print(paste0(disorder_now,' (exclude comorbidity): ', as.character(length(idx_disorder)), ' scans' ))
  comorbidity_dis <- demodf_s3[,disorder_now] * !demodf_s3$comorbidity_adhd_odd_anx
  num_consist <- table(comorbidity_dis,demodf_s3[,paste0(disorder_now, "_consistent")])[2,2]
  print(paste0("In the final sample, ",length(idx_disorder), " scans are diagnosed as ", disorder_now, " (exclude comorbidity odd & adhd), ", num_consist, " scans are consistent at two visits."))
  
  idx_disorder <- which(demodf_s3_cog[,disorder_now]==1 & demodf_s3_cog$comorbidity_adhd_odd_anx==FALSE)
  #print(paste0(disorder_now,' (exclude comorbidity): ', as.character(length(idx_disorder)), ' scans' ))
  comorbidity_dis <- demodf_s3_cog[,disorder_now] * !demodf_s3_cog$comorbidity_adhd_odd_anx
  num_consist <- table(comorbidity_dis,demodf_s3_cog[,paste0(disorder_now, "_consistent")])[2,2]
  print(paste0("In the cognition sample, ",length(idx_disorder), " scans are diagnosed as ", disorder_now, " (exclude comorbidity  odd & adhd), ", num_consist, " scans are consistent at two visits."))
}

# siteID from release 4
abcd_lt01 <- read.xlsx(paste0(datapath4, '/abcd_lt01.xlsx'))
demodf_s3 <- demodf_s3 %>% left_join(select(abcd_lt01, src_subject_id, eventname, site_id_l), join_by(src_subject_id, eventname))
demodf_s3 <- demodf_s3 %>% dplyr::rename(siteID4="site_id_l")
demodf_s3$siteID4 <- gsub("site0", "site", demodf_s3$siteID4)

write.csv(demodf_s3, paste0(info_path, '/DemodfScreenFinal.csv'), row.names = F)

```

```{r, p-factor, echo=FALSE}
# merge pfactor
pFactor.df <- read.csv("/Users/xuxiaoyu_work/Cuilab/open_dataset_information/ABCD/info/p.factor/BifactorP_baseTo3Year.csv")
pFactor.df$eventname <- tolower(pFactor.df$eventname)
pFactor.df <- dplyr::rename(pFactor.df, src_subject_id="srcsubjectid")
demodf_s4 <- base::merge(demodf_s3, pFactor.df, by=c("src_subject_id", "eventname"), all.x = T)

pFactor.WX0 <- read.csv("/Users/xuxiaoyu_work/Cuilab/open_dataset_information/ABCD/info/p.factor/4.pfactor_0y.csv")
pFactor.WX2 <- read.csv("/Users/xuxiaoyu_work/Cuilab/open_dataset_information/ABCD/info/p.factor/4.pfactor_2y.csv")
pFactor.WX0$scanID <- paste0("sub-", gsub("_", "", pFactor.WX0$ID), "_ses-baselineYear1Arm1")
pFactor.WX2$scanID <- paste0("sub-", gsub("_", "", pFactor.WX2$ID), "_ses-2YearFollowUpYArm1")
pFactor.WX <- rbind(pFactor.WX0, pFactor.WX2)

demodf_s4 <- demodf_s4 %>% left_join(select(pFactor.WX, c("general", "scanID")), by="scanID")
write.csv(demodf_s4, paste0(info_path, '/DemodfScreenFinal.csv'), row.names = F)
print(paste0(table(demodf_s4$eventname[!is.na(demodf_s4$general)])[2], " participants from baseline, ",  table(demodf_s4$eventname[!is.na(demodf_s4$general)])[1]," participants from 2-year FU in p-factor analyses."))

```

```{r SES, echo=FALSE}
abcd_y_lt <- read.csv(paste0(datapath, '/abcd-general/abcd_y_lt.csv'))
abcd_p_demo <- read.csv(paste0(datapath,'/abcd-general/abcd_p_demo.csv'))
demodf_s4 <- demodf_s4 %>% left_join(select(abcd_p_demo, c("src_subject_id", "eventname", "demo_comb_income_v2", "demo_roster_v2", "demo_comb_income_v2_l", "demo_roster_v2_l")), join_by("src_subject_id", "eventname"))
demodf_s4 <- demodf_s4 %>%
  group_by(src_subject_id) %>%
  mutate(demo_comb_income_v2 = coalesce(demo_comb_income_v2, demo_comb_income_v2_l)) %>%
  mutate(demo_roster_v2 = coalesce(demo_roster_v2, demo_roster_v2_l)) %>%
  ungroup()
demodf_s4$demo_comb_income_v2_l <- demodf_s4$demo_roster_v2_l <- NULL
demodf_s4$totalincome <- dplyr::recode(demodf_s4$demo_comb_income_v2, 
                                      `1` = 2500, `2` = 8500, `3` = 13999.5, `4` = 20499.5, `5` = 29999.5, `6` = 42499.5, `7` = 62499.5, `8` = 87499.5, `9` = 149999.5, `10` = 200000)
#1= Less than $5,000; 2=$5,000 through $11,999; 3=$12,000 through $15,999; 4=$16,000 through $24,999; 5=$25,000 through $34,999; 6=$35,000 through $49,999; 7=$50,000 through $74,999; 8= $75,000 through $99,999; 9=$100,000 through $199,999; 10=$200,000 and greater.
demodf_s4 <- demodf_s4 %>% left_join(select(abcd_y_lt, c("src_subject_id", "eventname", "interview_date")), join_by("src_subject_id", "eventname"))

demodf_s4$interviewYear<-substr(demodf_s4$interview_date,start=nchar(demodf_s4$interview_date)-3, stop=nchar(demodf_s4$interview_date))
demodf_s4$interviewYear=as.numeric(demodf_s4$interviewYear)
reference_data <- data.frame(familysize = rep(c(1:12),6),
                             povertyline = c(11880,16020,20160,24300,28440,32580,36730,40890,seq(from=45050, to=(45050+4160*3), by=4160), 12140,16460,20780,25100,29420,33740,38060,42380, 46700,51020,	55340,	59660, 12060,16240,20420,24600,28780,32960,37140,41320, 45500,	49680,	53860,	58040,12490,	16910,	21330,	25750,	30170,	34590,	39010,	43430,	47850,	52270,	56690,	61110,seq(from=12760, to=(12760+4480*11), by=4480), seq(from=12880, to=(12880+4540*11), by=4540)),Year = c(rep(2016,12),rep(2018,12), rep(2017,12), rep(2019,12),rep(2020,12),rep(2021,12)))
demodf_s4 <- merge(demodf_s4, reference_data,
                          by.x = c("interviewYear", "demo_roster_v2"),
                          by.y = c("Year", "familysize"), 
                          all.x = TRUE)
demodf_s4$income.adj<-demodf_s4$totalincome / demodf_s4$povertyline
summary(demodf_s4$income.adj)
demodf_s4$race_ethnicity <- factor(demodf_s4$race_ethnicity, levels=c(1:5), labels=c("White", "Black", "Hispanic", "Asian", "Other"))

write.csv(demodf_s4, paste0(info_path, '/DemodfScreenFinal.csv'), row.names = F)

```

## ICV
```{r ICV}
demodf_s4 <- read.csv(paste0(info_path, '/DemodfScreenFinal.csv'))
mri_y_smr_vol_aseg <- read.csv(paste0(datapath, "/imaging/mri_y_smr_vol_aseg.csv"))
demodf_s4 <- demodf_s4 %>% left_join(select(mri_y_smr_vol_aseg, c("src_subject_id", "eventname", "smri_vol_scs_intracranialv")), join_by("src_subject_id", "eventname"))
summary(demodf_s4$smri_vol_scs_intracranialv)
write.csv(demodf_s4, paste0(info_path, '/DemodfScreenFinal.csv'), row.names = F)

```
