---
title: "abcd_exclusion_demo_longitudinal"
author: "Xiaoyu xu"
date: "3/5/2024"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(tidyverse)
library(psych)
knitr::opts_chunk$set(echo = TRUE)
demo_path <- "/Volumes/exchange/abcd-data-release-5.0/abcd-data-release-5.1/core"
result_path <- "/Users/xuxiaoyu_work/Cuilab/open_dataset_information/ABCD/info"
```

# Minimal exclusion criteria

Across recruitment sites, inclusion criteria consisted of being in the required age range and able to provide informed consent (parents) and assent (child). Exclusions were minimal and were limited to lack of English language proficiency in the children, the presence of severe sensory, intellectual, medical or neurological issues that would impact the validity of collected data or the child’s ability to comply with the protocol, and contraindications to MRI scanning ([Garavan et al., 2018](https://www.sciencedirect.com/science/article/pii/S1878929317301809)). Parents must be fluent in either English or Spanish.

## 1) Language
1 = Poor; 2 = Fair; 3 = Good; 4 = Excellent; 999-Don’t know; 777-Refused
### 1.1) not fluent in English
```{r}
demo_data <- read.csv(paste0(demo_path,'/culture-environment/ce_y_acc.csv'))
demo_data <- demo_data[which(demo_data$eventname=='baseline_year_1_arm_1' | demo_data$eventname=='2_year_follow_up_y_arm_1'),]

idx_language_y <- with(demo_data, accult_q1_y==1)
idx_language_y[is.na(idx_language_y)] <- FALSE

demo_data$eventname2 <- gsub("baseline_year_1_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_language_y <- demo_data$scanID[idx_language_y]
num_language_y <- length(sub_language_y)
print(paste0('Not fluent in English: exclude ', as.character(num_language_y), ' participants' ))
```

### 1.2) parents not fluent in English or Spanish

```{r}
demo_data <- read.csv(paste0(demo_path,'/culture-environment/ce_p_acc.csv'))
demo_data <- demo_data[which(demo_data$eventname=='baseline_year_1_arm_1' | demo_data$eventname=='2_year_follow_up_y_arm_1'),]

idx_english_ok <- with(demo_data, accult_q1_p>1)
idx_english_ok[is.na(idx_english_ok)] <- FALSE

idx_spanish_ok <- with(demo_data, accult_q3_dropdwn_p==47 & accult_q3b_p>1)
idx_spanish_ok[is.na(idx_spanish_ok)] <- FALSE

idx_language_p <- !(idx_english_ok | idx_spanish_ok)

demo_data$eventname2 <- gsub("baseline_year_1_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_language_p <- demo_data$scanID[idx_language_p]
num_language_p <- length(sub_language_p)
print(paste0('Parents not fluent in English or Spanish: exclude ', as.character(num_language_p), ' participants' ))
```

### 2) Traumatic brain injury history
Worst Injury Overall:

1, Improbable TBI (no TBI or TBI w/o LOC or memory loss);

2, Possible mild TBI (TBI w/o LOC but memory loss);

3, Mild TBI (TBI w/LOC ≤ 30 min);

4, Moderate TBI (TBI w/LOC 30 min - 24 hrs);

5, Severe TBI (TBI w/ LOC ≥ 24 hrs)
```{r}
demo_data <- read.csv(paste0(demo_path,'/physical-health/ph_p_otbi.csv'))
demo_data <- demo_data[which(demo_data$eventname=='baseline_year_1_arm_1' | demo_data$eventname=='2_year_follow_up_y_arm_1'),]

idx_tbi <- with(demo_data, tbi_ss_worst_overall>2 | tbi_ss_worst_overall_l>2)
idx_tbi[is.na(idx_tbi)] <- FALSE

demo_data$eventname2 <- gsub("baseline_year_1_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_tbi <- demo_data$scanID[idx_tbi]
num_tbi <- length(sub_tbi)
print(paste0('Traumatic brain injury history: exclude ', as.character(num_tbi), ' participants' ))

```

### 3) Sensory issues
```{r}
demo_data <- read.csv(paste0(demo_path,'/abcd-general/abcd_p_screen.csv'))

idx_sensory <- with(demo_data, scrn_percept==1 & scrn_percept_corr==0)
idx_sensory[is.na(idx_sensory)] <- FALSE

demo_data$eventname2 <- gsub("screener_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_sensory <- demo_data$scanID[idx_sensory]
num_sensory <- length(sub_sensory)
print(paste0('Sensory issues: exclude ', as.character(num_sensory), ' participants' ))
```

### 4) Intellectual disability
```{r}
demo_data <- read.csv(paste0(demo_path,'/abcd-general/abcd_p_screen.csv'))

idx_intel <- with(demo_data, scrn_intdisab==1)
idx_intel[is.na(idx_intel)] <- FALSE

demo_data$eventname2 <- gsub("screener_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_intel <- demo_data$scanID[idx_intel]
num_intel <- length(sub_intel)
print(paste0('Intellectual disability: exclude ', as.character(sum(idx_intel)), ' participants at baseline' ))

```

### 5) Major medical or neurological conditions
```{r}
demo_data <- read.csv(paste0(demo_path,'/abcd-general/abcd_p_screen.csv'))

idx_medical <- with(demo_data, scrn_cpalsy==0 | scrn_tumor==0 | scrn_stroke==0 | scrn_aneurysm==0 | scrn_hemorrhage==0 |
                      scrn_hemotoma==0 | scrn_medcond_other==1)
idx_medical[is.na(idx_medical)] <- FALSE

demo_data$eventname2 <- gsub("screener_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_medical <- demo_data$scanID[idx_medical]
num_medical <- length(sub_medical)
print(paste0('Major medical or neurological conditions: exclude ', as.character(num_medical), ' participants at baseline' ))

```

### 6) Contraindications to MRI scanning
```{r}
demo_data <- read.csv(paste0(demo_path,'/abcd-general/abcd_p_screen.csv'))
idx_contra <- with(demo_data, scrn_met_obj==1 | scrn_nr_piercing==1 | scrn_weight_mri==1 | 
                      scrn_met_injremove==0 | scrn_bracescallback==0)
idx_contra[is.na(idx_contra)] <- FALSE

demo_data$eventname2 <- gsub("screener_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_contra <- demo_data$scanID[idx_contra]
num_contra <- length(sub_contra)
print(paste0('Contraindications to MRI scanning: exclude ', as.character(num_contra), ' participants at baseline' ))

```

### 7) Premature & low birth weight (Optional)
```{r}
demo_data <- read.csv(paste0(demo_path,'/physical-health/ph_p_dhx.csv'))
demo_data <- demo_data[demo_data$eventname=='baseline_year_1_arm_1',]

idx_birth <- with(demo_data, devhx_12a_p==1)
idx_birth[is.na(idx_birth)] <- FALSE

# 1 lbs = 16 oz, low birth weight = 5 lbs,8 oz
demo_data$birth_weight <- demo_data$birth_weight_lbs + demo_data$birth_weight_oz/16
idx_weight <- with(demo_data, birth_weight<5.5)
idx_weight[is.na(idx_weight)] <- FALSE

idx_premature <- idx_birth | idx_weight

demo_data$eventname2 <- gsub("baseline_year_1_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)

sub_premature <- demo_data$scanID[idx_premature]
num_premature <- length(sub_premature)
print(paste0('Premature & low birth weight: exclude ', as.character(num_premature), ' participants at baseline' ))

```

## Summary
```{r, include=TRUE}
print(paste0('Not fluent in English: exclude ', as.character(num_language_y), ' participants' ))
print(paste0('Parents not fluent in English or Spanish: exclude ', as.character(num_language_p), ' participants' ))
print(paste0('Traumatic brain injury history: exclude ', as.character(num_tbi), ' participants' ))
print(paste0('Sensory issues: exclude ', as.character(num_sensory), ' participants' ))
print(paste0('Intellectual disability: exclude ', as.character(sum(idx_intel)), ' participants at baseline' ))
print(paste0('Major medical or neurological conditions: exclude ', as.character(num_medical), ' participants at baseline' ))
print(paste0('Contraindications to MRI scanning: exclude ', as.character(num_contra), ' participants at baseline' ))
print(paste0('Premature & low birth weight: exclude ', as.character(num_premature), ' participants at baseline' ))
print(paste0('In total: exclude ', sum(num_language_y, num_language_p, num_tbi, num_sensory, sum(idx_intel), num_medical, num_contra, num_premature), ' participants'))

```
### Sex
Exclude participants with invalid Sex (>2,NA).

(1 = Male Masculino; 2 = Female Femenino; 3 = Intersex-Male Entre sexo-masculino; 4 = Intersex-Female Entre sexo-femenino; 777 = Refuse to answer; 999 = Don’t know)

```{r}
demo_data <- read.csv(paste0(demo_path,'/abcd-general/abcd_p_demo.csv'))
demo_data <- demo_data[which(demo_data$eventname=='baseline_year_1_arm_1' | demo_data$eventname=='2_year_follow_up_y_arm_1'),]

demo_data$demo_brthdat_v2[which(demo_data$eventname=='2_year_follow_up_y_arm_1')] <- demo_data$demo_brthdat_v2_l[which(demo_data$eventname=='2_year_follow_up_y_arm_1')]

demo_data$eventname2 <- gsub("baseline_year_1_arm_1", "baselineYear1Arm1", demo_data$eventname)
demo_data$eventname2 <- gsub("2_year_follow_up_y_arm_1", "2YearFollowUpYArm1", demo_data$eventname2)
demo_data$scanID <- paste0(gsub("NDAR_", "sub-NDAR", demo_data$src_subject_id),"_ses-", demo_data$eventname2)
demo_data <- demo_data %>% group_by(src_subject_id) %>% 
  mutate(demo_sex_v2 = zoo::na.locf(demo_sex_v2)) %>%
  mutate(demo_sex_v2 = zoo::na.locf(demo_sex_v2, fromLast = TRUE, na.rm = FALSE))

idx_sex <- with(demo_data, demo_sex_v2>2 | is.na(demo_sex_v2))
idx_sex[is.na(idx_sex)] <- TRUE
sub_sex <- demo_data$scanID[idx_sex]
print(paste0('Invalid sex: exclude ', sum(idx_sex), ' participants' ))

```
## merge data
Write the criteria out.
```{r , include=FALSE}
demo_sublist <- demo_data[,c("race_ethnicity", "scanID", "src_subject_id", "eventname")]
## merge data
demo_sublist$subID <- str_split_i(demo_sublist$scanID, "_ses-", 1)
demo_sublist$if_language_y <- demo_sublist$scanID %in% sub_language_y
demo_sublist$if_language_p <- demo_sublist$scanID %in% sub_language_p
demo_sublist$if_TBI <- demo_sublist$scanID %in% sub_tbi

sub_sensory <- str_split_i(sub_sensory, "_ses-", 1)
demo_sublist$if_sensory <- demo_sublist$subID %in% sub_sensory
sub_intel <- str_split_i(sub_intel, "_ses-", 1)
demo_sublist$if_intel <- demo_sublist$subID %in% sub_intel
sub_medical <- str_split_i(sub_medical, "_ses-", 1)
demo_sublist$if_medical <- demo_sublist$subID %in% sub_medical
sub_contra <- str_split_i(sub_contra, "_ses-", 1)
demo_sublist$if_contra <- demo_sublist$subID %in% sub_contra
sub_premature <- str_split_i(sub_premature, "_ses-", 1)
demo_sublist$if_premature <- demo_sublist$subID %in% sub_premature
demo_sublist$if_gender <- demo_sublist$scanID %in% sub_sex

write.csv(demo_sublist, paste0(result_path, "/demo_sublist_criteria.csv"), row.names = F)
```

